### 2023-04-11 paper/localpsf.tex
edits during zoom meeting with Noemi and Tucker

We use our method to approximate the Newton (or Gauss-Newton) Hessian
in an inverse problem governed by a PDE which models steady state ice
sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
and flow of a contaminant~\cite{PetraStadler11}
(Section~\ref{sec:adv}). To reconstruct the unknown parameters, we
formulate an inverse problem governed by the corresponding PDEs. In
particular this problem is formulated as a nonlinear least squares
optimization problem, whose cost function consists of a misfit term
(between the observations and model output) and a bi-Laplacian
regularization term following~\cite{VillaPetraGhattas21}. The
regularization is centered at a constant function $q_0(x)$. To
mitigate boundary effects
%For stokes q_0(x) = 10.5$%
we use a constant coefficient Robin boundary condition as
in~\cite{Roininen14}. The parameters for the regularization are chosen
so that the Green's function of the Hessian of the regularization has
a characteristic length of $0.25$ of the domain radius. For the
specific setup, we refer the reader to~\cite[Section
  2.2]{VillaPetraGhattas21}. For both examples, we choose the
regularization parameter, $\gamma$, that controls the overall strength
of the regularization using Morozov's discrepancy
principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
all the numerical results except for the studies where study how
changing $\gamma$ impacts the effectiveness of the proposed
preconditioner.

We solve the inverse problem with an inexact Newton preconditioned
conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
$\bm{\searchdir}_k$, are obtained by solving 
%
\begin{equation}
	\label{eq:newton_system}
	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
\end{equation}
%
wherein we choose the initial guess as the discretization of the constant
function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
the discretized gradient, Hessian, and Gauss-Newton Hessian of the
inverse problem cost function, respectively.
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
To ensure positive definiteness of the Hessian we use
$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
subsequent iterations. The Newton iterations are terminated when
$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 batches. These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively. The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix~\ref{sec:make_spd}, with $l=2$. We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal basal sliding parameter. Our results show that our preconditioners are also good preconditioners for $\bm{H}$. The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
%
%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
%$
%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
%$
%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
%
We note that the Hessian of the regularization is a combination of
stiffness and mass matrices, which are sparse. Therefore, we form
H-matrix representations of these matrices and combine them using
standard sparse H-matrix techniques.  We factor the overall
Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$,
using fast H-matrix methods, then use the factorization as a
preconditioner.

@@

\begin{figure}
	\begin{subfigure}{0.59\textwidth}
		\begin{center}
			\begin{tikzpicture}[baseline, trim axis right]
			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
			ymode=log,
			xmin=-50,
			xmax=1450,
			ymax=1.e4,
			xlabel={$k$, generalized eigenvalue \#},
			ylabel={$\lambda_{k}$},
			xtick={0, 300, 600, 900, 1200},
			%legend pos=north east,
			legend style={font=\small},
			ylabel near ticks, ylabel shift={-6pt},
			axis x line*=bottom,
			axis y line*=left,
			legend style={at={(0.95,1)},anchor=north east},
			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
			]
			\addlegendentry{REG}
			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
			\end{axis}
			\end{tikzpicture}
		\end{center}
		\caption{}
		\label{fig:generalized_eigenvalues}
	\end{subfigure}
	\begin{subfigure}{0.39\textwidth}
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{4pt}
			\renewcommand{\arraystretch}{1.25}
			\begin{tabular}{c| c c}
				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$ &	$\kappa(\bm{\preconditioner}^{-1} \bm{H}_{\text{gn}})$ \\
				\hline 
				REG        & 7.02e+3 & 7.02e+3 \\
				NONE       & 8.71e+2 & 9.22e+2 \\
				PSF $(1)$  & 1.75e+1 & 1.76e+1 \\
				PSF $(5)$  & 1.39e+1 & 1.39e+1 \\
				PSF $(25)$ & 1.03e+1 & 1.06e+1
			\end{tabular}
			\endgroup
		\end{center}
		\caption{}
		\label{fig:condition_number}
	\end{subfigure}
	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$. Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG). Right (\ref{fig:condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE). All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle. COND(H) LABELS!}
	\label{fig:stokes_spectrum}
\end{figure}

In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
$
	\bm{H} \bm{u} = \lambda \bm{\preconditioner} \bm{u}.
$
Here $\bm{H}$ is the Hessian evaluated at the reconstructed parameter $\bm{\basalfriction}$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle. The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\bm{\basalfriction}$, or the regularization Hessian. The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering. The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues. In Figure \ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$. The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers. Regularization preconditioning and no preconditioning yield larger condition numbers. The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.

@@

@@

%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\D)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
%
%$$
%\begin{split}
%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \D\times(0,T),\\
%u(\cdot, 0) &= \ipar & \quad \text{in } \D,\\
%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial\D \times (0,T).
%\end{split}
%$$
%
%Here, $\D \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
%time. The velocity field
%$\vec{v}$ is computed by solving the following steady-state
%Navier-Stokes equation with the side walls driving the flow:
%
%$$
%\begin{aligned}
%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\D,\\
%\nabla \cdot {\vec v} &= 0 &&\text{ in }\D,\\
%{\vec v} &= {\vec g} &&\text{ on } \partial\D.
%\end{aligned}
%$$
%
%Here, $q$ is pressure, $\text{Re}$ is the Reynolds number. The Dirichlet boundary data
%${\vec g} \in \R^d$ is given by 
%${\vec g} = {\vec e}_2$ on the left wall of the domain, 
%${\vec g}=-{\vec e}_2$ on the right wall,  and ${\vec g} = {\vec 0}$ everywhere else.

@@

\begin{figure}
	\begin{subfigure}{0.59\textwidth}
		\begin{center}
			\begin{tikzpicture}[baseline, trim axis right]
			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
			ymode=log,
			xmin=-50,
			xmax=1450,
			ymax=1.e4,
			xlabel={$k$, generalized eigenvalue \#},
			ylabel={$\lambda_{k}$},
			xtick={0, 300, 600, 900, 1200},
			%legend pos=north east,
			legend style={font=\small},
			ylabel near ticks, ylabel shift={-6pt},
			axis x line*=bottom,
			axis y line*=left,
			legend style={at={(0.95,1)},anchor=north east},
			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
			]
			\addlegendentry{REG}
			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
			\end{axis}
			\end{tikzpicture}
		\end{center}
		\caption{}
		\label{fig:adv_generalized_eigenvalues}
	\end{subfigure}
	\begin{subfigure}{0.39\textwidth}
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{4pt}
			\renewcommand{\arraystretch}{1.25}
			\begin{tabular}{c| c c}
				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$\\
				\hline 
				REG        & 7.02e+3 \\
				NONE       & 8.71e+2 \\
				PSF $(5)$  & 1.75e+1 \\
				PSF $(15)$  & 1.39e+1 \\
				PSF $(30)$ & 1.03e+1
			\end{tabular}
			\endgroup
		\end{center}
		\caption{}
		\label{fig:adv_condition_number}
	\end{subfigure}
	\caption{Left (\ref{fig:adv_generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$. Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 5, 15, and 30 batches (PSF (5), PSF (15), and PSF (30), respectively), or the regularization Hessian (REG). Right (\ref{fig:adv_condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE). COND(H) LABELS!}
	\label{fig:adv_spectrum}
\end{figure}

>>>>

We use our method to approximate the Newton (or Gauss-Newton) Hessian
in inverse problems governed by PDEs which model steady state ice
sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
and flow of a contaminant~\cite{PetraStadler11}
(Section~\ref{sec:adv}). 
To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
optimization problems, whose cost functions consist of a misfit term
(between the observations and model output) and a bi-Laplacian
regularization term following~\cite{VillaPetraGhattas21}. The
regularization is centered at a constant function $q_0(x)$. To
mitigate boundary effects
%For stokes q_0(x) = 10.5$%
we use a constant coefficient Robin boundary condition as
in~\cite{Roininen14}. The parameters for the regularization are chosen
so that the Green's function of the Hessian of the regularization has
a characteristic length of $0.25$ (NOTE: CHECK THIS FOR BOTH PROBLEMS) of the domain radius. For the
specific setup, we refer the reader to~\cite[Section
  2.2]{VillaPetraGhattas21}. For both examples, we choose the
regularization parameter, $\gamma$, that controls the overall strength
of the regularization using Morozov's discrepancy
principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
all the numerical results except for the studies where we study how
changing $\gamma$ impacts the effectiveness of the proposed
preconditioner.

We solve the ice sheet inverse problem with an inexact Newton preconditioned
conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
$\bm{\searchdir}_k$, are obtained by solving 
%
\begin{equation}
	\label{eq:newton_system}
	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
\end{equation}
%
wherein we choose the initial guess as the discretization of the constant
function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
the discretized gradient, Hessian, and Gauss-Newton Hessian of the
inverse problem cost function, respectively.
%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
To ensure positive definiteness of the Hessian we use
$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
subsequent iterations. The Newton iterations are terminated when
$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$. The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration. In this case the Newton linear system, \eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section \ref{sec:adv}.

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit). The 
%number of impulse response batches used to build these 
approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation. The Hessian of the regularization term is a combination of
stiffness and mass matrices, which are sparse. Therefore, we form
H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
standard sparse H-matrix techniques. H-matrix approximations of the Gauss-Newton Hessian, 
$$\bm{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic. The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Appendix~\ref{sec:make_spd}. We factor $\bm{\preconditioner}$,
using fast H-matrix methods, then use the factorization as a
preconditioner. We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel. Our results show that $\widetilde{\mathbf{H}}$ is a good preconditioner for both $\bm{H}_\text{gn}$ and $\bm{H}$. 
%The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
%
%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
%$
%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
%$
%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
%

@@

\begin{figure}
	\begin{subfigure}{0.59\textwidth}
		\begin{center}
			\begin{tikzpicture}[baseline, trim axis right]
			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
			ymode=log,
			xmin=-50,
			xmax=1450,
			ymax=1.e4,
			xlabel={$k$, generalized eigenvalue \#},
			ylabel={$\lambda_{k}$},
			xtick={0, 300, 600, 900, 1200},
			%legend pos=north east,
			legend style={font=\small},
			ylabel near ticks, ylabel shift={-6pt},
			axis x line*=bottom,
			axis y line*=left,
			legend style={at={(0.95,1)},anchor=north east},
			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
			]
			\addlegendentry{REG}
			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
			\end{axis}
			\end{tikzpicture}
		\end{center}
		\caption{}
		\label{fig:generalized_eigenvalues}
	\end{subfigure}
	\begin{subfigure}{0.39\textwidth}
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{4pt}
			\renewcommand{\arraystretch}{1.25}
			\begin{tabular}{c| c c}
				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$ &	$\kappa(\bm{\preconditioner}^{-1} \bm{H}_{\text{gn}})$ \\
				\hline 
				REG        & 7.02e+3 & 7.02e+3 \\
				NONE       & 8.71e+2 & 9.22e+2 \\
				PSF $(1)$  & 1.75e+1 & 1.76e+1 \\
				PSF $(5)$  & 1.39e+1 & 1.39e+1 \\
				PSF $(25)$ & 1.03e+1 & 1.06e+1
			\end{tabular}
			\endgroup
		\end{center}
		\caption{}
		\label{fig:condition_number}
	\end{subfigure}
	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \genericvec_k = \lambda_k \bm{\preconditioner} \genericvec_k$. Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG). Right (\ref{fig:condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE). All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle. COND(H) LABELS!}
	\label{fig:stokes_spectrum}
\end{figure}

In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
$
	\bm{H} \genericvec = \lambda \bm{\preconditioner} \genericvec.
$
Here $\bm{H}$ is the Hessian evaluated at the reconstructed parameter $\bm{\basalfriction}$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle. The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\bm{\basalfriction}$, or the regularization Hessian. The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering. The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues. In Figure \ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$. The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers. Regularization preconditioning and no preconditioning yield larger condition numbers. The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.

@@

Here we consider a time-dependent advection-diffusion equation for
which we seek to infer the unknown spatially varying initial condition, $\ipar$, from observations of the state  
at a final time. This PDE
models diffusive transport in a domain $\advdomain \subset \R^d$, which is
depicted in Figure~\ref{fig:adv_inversion1}. In this case, the state, $\concentration(x,t)$, could be interpreted as the concentration of a contaminant. The problem description
below closely follows the one in~\cite{PetraStadler11}. The domain
boundaries $\partial \advdomain$ include the outer boundaries as well as the
internal boundaries of the rectangles, which represent buildings.  The
parameter-to-observable map $\iFF$ in this case maps an initial
condition $\ipar \in L^2(\advdomain)$ to observations of the concentration at a final time, $\concentration(x, T)$,  through solution of the advection-diffusion
equation given by
\begin{equation}\label{eq:ad}
  \begin{aligned}
    \concentration_t - \kappa\Delta \concentration + v\cdot\nabla \concentration &= 0 & \quad&\text{in
    }\advdomain\times (0,T), \\
    \concentration(\cdot, 0) &= \ipar  &&\text{in } \advdomain , \\
    \kappa\nabla \concentration\cdot \nu &= 0 &&\text{on } \partial \advdomain \times (0,T).
  \end{aligned}
\end{equation}
Here, $\kappa > 0$ is a diffusivity coefficient, $\nu$ is the boundary unit normal vector, and $T > 0$ is the
final time.  The velocity field $v:\advdomain \rightarrow \mathbb{R}^d$, shown in
Figure~\ref{fig:adv_inversion1}, is computed by solving the
steady-state Navier-Stokes equations for a two dimensional flow with
Reynolds number 50 and boundary conditions as in
Figure~\ref{fig:adv_inversion} (left); see \cite{PetraStadler11} for
details.  The time evolution of the state variable $\concentration$ from a given
initial condition $\ipar$ is illustrated in
Figure~\ref{fig:adv_inversion1}.

@@

%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\advdomain)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
%
%$$
%\begin{split}
%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \advdomain\times(0,T),\\
%u(\cdot, 0) &= \ipar & \quad \text{in } \advdomain,\\
%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial \advdomain \times (0,T).
%\end{split}
%$$
%
%Here, $\advdomain \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
%time. The velocity field
%$\vec{v}$ is computed by solving the following steady-state
%Navier-Stokes equation with the side walls driving the flow:
%
%$$
%\begin{aligned}
%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\advdomain,\\
%\nabla \cdot {\vec v} &= 0 &&\text{ in }\advdomain,\\
%{\vec v} &= {\vec g} &&\text{ on } \partial \advdomain.
%\end{aligned}
%$$
%
%Here, $q$ is pressure, $\text{Re}$ is the Reynolds number. The Dirichlet boundary data
%${\vec g} \in \R^d$ is given by 
%${\vec g} = {\vec e}_2$ on the left wall of the domain, 
%${\vec g}=-{\vec e}_2$ on the right wall,  and ${\vec g} = {\vec 0}$ everywhere else.

@@

\begin{figure}
	\begin{subfigure}{0.59\textwidth}
		\begin{center}
			\begin{tikzpicture}[baseline, trim axis right]
			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
			ymode=log,
			xmin=-50,
			xmax=1450,
			ymax=1.e4,
			xlabel={$k$, generalized eigenvalue \#},
			ylabel={$\lambda_{k}$},
			xtick={0, 300, 600, 900, 1200},
			%legend pos=north east,
			legend style={font=\small},
			ylabel near ticks, ylabel shift={-6pt},
			axis x line*=bottom,
			axis y line*=left,
			legend style={at={(0.95,1)},anchor=north east},
			title={Generalized eigenvalues, $\bm{H} \genericvec_{k}=\lambda_{k} \bm{\preconditioner} \genericvec_{k}$}
			]
			\addlegendentry{REG}
			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (1)}
			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (5)}
			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
			\addlegendentry{PSF (25)}
			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
			\end{axis}
			\end{tikzpicture}
		\end{center}
		\caption{}
		\label{fig:adv_generalized_eigenvalues}
	\end{subfigure}
	\begin{subfigure}{0.39\textwidth}
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{4pt}
			\renewcommand{\arraystretch}{1.25}
			\begin{tabular}{c| c c}
				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$\\
				\hline 
				REG        & 7.02e+3 \\
				NONE       & 8.71e+2 \\
				PSF $(5)$  & 1.75e+1 \\
				PSF $(15)$  & 1.39e+1 \\
				PSF $(30)$ & 1.03e+1
			\end{tabular}
			\endgroup
		\end{center}
		\caption{}
		\label{fig:adv_condition_number}
	\end{subfigure}
	\caption{Left (\ref{fig:adv_generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$. Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 5, 15, and 30 batches (PSF (5), PSF (15), and PSF (30), respectively), or the regularization Hessian (REG). Right (\ref{fig:adv_condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE). COND(H) LABELS!}
	\label{fig:adv_spectrum}
\end{figure}
